---
title: "04 Check Clone Genotype Cell Metadata Consistency"
output: html_document
---

```{r setup, message=FALSE}
library(parallel)
```

## Check

In v3, count metadata keeps barcodes with at least one RNA count, so feature
h5ad barcodes do not need to be a subset of count metadata. The required check
is narrower: count metadata cells with a nonempty clone or genotype annotation
must exist in the corresponding feature h5ad and have the same decoded feature
annotation.

The feature h5ad `top_feature_name` is not the comparison value. Decode
`top_feature_index` through `var$feature_name`.

| source | barcode | value |
|---|---|---|
| count metadata clone ID | cell_A | BC201125 |
| feature obs top_feature_index | cell_A | 201125 |
| feature obs top_feature_name | cell_A | BC201126 |
| feature var feature_name[201125] |  | BC201125 |

```{r verify}
cell_meta_dir = "/fh/fast/sun_w/yub/github/MSK_30polyKO/dracc_qc/outputs/01_extract_raw_count_cell_meta"

dracc_dir = Sys.getenv("DRACC_PROCESSED_DIR")
corrected_de_xm_dir = Sys.getenv("DRACC_DE_XM_DIR")

stage_order = c("ES", "DE", "DE_XM", "PP1", "PP2", "S5_1", "S5_2", "S6_1", "S6_2")
stage_suffix = c(
  ES = "es",
  DE = "de",
  DE_XM = "de_gemx",
  PP1 = "pp1",
  PP2 = "pp2",
  S5_1 = "s5_1",
  S5_2 = "s5_2",
  S6_1 = "s6_1",
  S6_2 = "s6_2"
)

# Verify one metadata call column against one feature h5ad.
check_feature_calls <- function(meta, feature_path, call_col) {
  h5ad = anndata::read_h5ad(feature_path, backed = "r")
  obs = as.data.frame(h5ad$obs, stringsAsFactors = FALSE)
  feature_names = as.character(h5ad$var$feature_name)

  stopifnot("top_feature_index" %in% colnames(obs))
  stopifnot(!anyDuplicated(rownames(obs)))
  stopifnot(!anyNA(meta[[call_col]]))

  meta_call = as.character(meta[[call_col]])
  called = meta_call != ""
  feature_match = match(meta$dracc_barcode[called], rownames(obs))
  stopifnot(!anyNA(feature_match))

  feature_index = as.integer(as.character(obs$top_feature_index[feature_match]))
  decoded_call = rep(NA_character_, length(feature_index))
  valid_index = !is.na(feature_index) & feature_index > 0 & feature_index <= length(feature_names)
  decoded_call[valid_index] = feature_names[feature_index[valid_index]]
  stopifnot(!anyNA(decoded_call))
  stopifnot(identical(meta_call[called], decoded_call))

  TRUE
}

# Verify one stage.
check_stage <- function(stage_id) {
  stage_dir = if (stage_id == "DE_XM") {
    corrected_de_xm_dir
  } else {
    file.path(dracc_dir, paste0("30_KO_", stage_id))
  }

  meta_path = file.path(cell_meta_dir, paste0(stage_id, ".rds"))
  clone_col = paste0("Custom_larry_", stage_suffix[[stage_id]], "__feature_call")
  genotype_col = paste0("CRISPR_Guide_Capture_grna_", stage_suffix[[stage_id]], "__feature_call")

  stopifnot(file.exists(meta_path))
  meta = readRDS(meta_path)
  stopifnot(all(c("stage_id", "dracc_barcode", clone_col, genotype_col) %in% colnames(meta)))
  stopifnot(all(meta$stage_id == stage_id), !anyNA(meta$dracc_barcode))

  check_feature_calls(meta, file.path(stage_dir, "features", "larry", "features.h5ad"), clone_col)
  check_feature_calls(meta, file.path(stage_dir, "features", "gene", "features.h5ad"), genotype_col)

  TRUE
}

stage_checks = parallel::mclapply(
  stage_order,
  check_stage,
  mc.cores = min(length(stage_order), parallel::detectCores())
)

failed = vapply(stage_checks, inherits, logical(1), "try-error")
if (any(failed)) {
  stop(stage_order[which(failed)[1]], ": ", stage_checks[[which(failed)[1]]])
}

stopifnot(all(unlist(stage_checks)))
cat("All clone and genotype metadata calls are consistent with feature h5ad annotations.\n")
```
