---
title: "MSK Workflow Validation"
output:
  html_document:
    toc: true
    toc_float: true
---

This audit checks the supplied QC labels and reconstructs the final MSK cell
set from the transcriptome-QC and barcode-assignment metadata.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

source_root = "/fh/fast/sun_w/yub/github/MSK_30polyKO/external_sources/msk/30KO_scRNAseq"
tx_dir = file.path(source_root, "3.TX_QC")
bc_dir = file.path(source_root, "2.demulti_BC")
nan_final_path = file.path(source_root, "df.meta.rds")
canonical_final_path = "/fh/fast/sun_w/yub/data/msk_30ko/raw/df.meta.rds"
```

## Inputs

The validation reads Nan's supplied metadata directly and compares its final
object with the canonical MSK metadata.

```{r load-inputs}
# 1. Load and check the supplied inputs
tx_paths = sort(list.files(tx_dir, pattern = "rds$", full.names = TRUE))
bc_paths = sort(list.files(bc_dir, pattern = "rds$", full.names = TRUE))

stopifnot(
  length(tx_paths) == 9L,
  length(bc_paths) == 9L,
  file.exists(nan_final_path),
  file.exists(canonical_final_path)
)

nan_final = readRDS(nan_final_path)
final_meta = readRDS(canonical_final_path)

input_summary = data.frame(
  input = c(
    "Transcriptome-QC metadata files",
    "Barcode-assignment metadata files",
    "Nan final metadata cells",
    "Canonical final metadata cells"
  ),
  value = c(
    length(tx_paths),
    length(bc_paths),
    nrow(nan_final),
    nrow(final_meta)
  )
)

knitr::kable(input_summary, row.names = FALSE)

final_objects_identical = identical(nan_final, final_meta)
knitr::kable(
  data.frame(
    check = "Nan and canonical final R objects are identical",
    passed = final_objects_identical
  ),
  row.names = FALSE
)

stopifnot(final_objects_identical)
```

## Transcriptome QC

The expected labels are reconstructed from the four documented transcriptome
cutoffs. Every supplied label and the combined `Qualified` value must agree.

```{r validate-transcriptome-qc}
# 2. Recalculate the documented transcriptome-QC labels
tx_stages = sub(
  ".seu.QC.metadata.rds",
  "",
  sub("df.", "", basename(tx_paths), fixed = TRUE),
  fixed = TRUE
)

stopifnot(
  !anyDuplicated(tx_stages),
  !anyDuplicated(tx_paths)
)

tx_qc_results = lapply(seq_along(tx_paths), \(i) {
  x = readRDS(tx_paths[[i]])

  expected_qc1 = !is.na(x$nCount_RNA) & x$nCount_RNA >= 200
  expected_qc2 = (
    !is.na(x$scDblFinder.class) &
      x$scDblFinder.class == "singlet"
  )
  expected_qc3 = !is.na(x$nFeature_RNA) & x$nFeature_RNA >= 100
  expected_qc4 = !is.na(x$percent_mito) & x$percent_mito <= 20
  expected_qualified = (
    expected_qc1 &
      expected_qc2 &
      expected_qc3 &
      expected_qc4
  )

  expected_qc1_label = ifelse(
    expected_qc1,
    "nCount_RNA>=200",
    "nCount_RNA<200"
  )
  expected_qc2_label = ifelse(expected_qc2, "singlet", "doublet")
  expected_qc3_label = ifelse(
    expected_qc3,
    "nFeature_RNA>=100",
    "nFeature_RNA<100"
  )
  expected_qc4_label = ifelse(
    expected_qc4,
    "percent_mito<=20",
    "percent_mito>20"
  )

  data.frame(
    source_stage = tx_stages[[i]],
    input_cells = nrow(x),
    qualified_cells = sum(x$Qualified == "Yes", na.rm = TRUE),
    qc1_mismatch = sum(
      is.na(x$QC1_nCount_RNA) |
        as.character(x$QC1_nCount_RNA) != expected_qc1_label
    ),
    qc2_mismatch = sum(
      is.na(x$QC2_Doublets) |
        as.character(x$QC2_Doublets) != expected_qc2_label
    ),
    qc3_mismatch = sum(
      is.na(x$QC3_nFeature_RNA) |
        as.character(x$QC3_nFeature_RNA) != expected_qc3_label
    ),
    qc4_mismatch = sum(
      is.na(x$QC4_mito) |
        as.character(x$QC4_mito) != expected_qc4_label
    ),
    qualified_mismatch = sum(
      is.na(x$Qualified) |
        (as.character(x$Qualified) == "Yes") != expected_qualified
    )
  )
})

tx_qc_results = do.call(rbind, tx_qc_results)
knitr::kable(tx_qc_results, row.names = FALSE)

stopifnot(
  all(tx_qc_results$qc1_mismatch == 0L),
  all(tx_qc_results$qc2_mismatch == 0L),
  all(tx_qc_results$qc3_mismatch == 0L),
  all(tx_qc_results$qc4_mismatch == 0L),
  all(tx_qc_results$qualified_mismatch == 0L)
)
```

## Barcode Categories

The supplied tables do not contain the second-largest UMI, so the
second/maximum ratio cannot be recalculated. The available top-UMI values are
checked against the Type-1 minimum of five UMIs.

```{r validate-barcode-categories}
# 3. Check the recoverable part of the barcode category definitions
bc_stages = sub(
  ".rds",
  "",
  sub(
    "df.all.larryBC.geneBC.",
    "",
    basename(bc_paths),
    fixed = TRUE
  ),
  fixed = TRUE
)

stopifnot(
  setequal(tx_stages, bc_stages),
  !anyDuplicated(bc_stages),
  !anyDuplicated(bc_paths)
)

type1 = "Type 1: Max >=5 & 2nd/Max <= 0.5"
type3 = "Type 3: Max < 5"

bc_category_results = lapply(seq_along(bc_paths), \(i) {
  x = readRDS(bc_paths[[i]])
  larry_umi = suppressWarnings(as.numeric(as.character(x$larryUMI)))
  gene_umi = suppressWarnings(as.numeric(as.character(x$geneBCUMI)))
  larry_type1 = x$larryBCCategory %in% type1
  gene_type1 = x$geneBCCategory %in% type1

  data.frame(
    source_stage = bc_stages[[i]],
    input_cells = nrow(x),
    larry_type1_cells = sum(larry_type1),
    larry_type1_umi_failure = sum(
      larry_type1 & (is.na(larry_umi) | larry_umi < 5)
    ),
    gene_type1_cells = sum(gene_type1),
    gene_type1_umi_failure = sum(
      gene_type1 & (is.na(gene_umi) | gene_umi < 5)
    )
  )
})

bc_category_results = do.call(rbind, bc_category_results)
knitr::kable(bc_category_results, row.names = FALSE)

stopifnot(
  all(bc_category_results$larry_type1_umi_failure == 0L),
  all(bc_category_results$gene_type1_umi_failure == 0L)
)
```

## Final Cell Set

The final selection is reconstructed by joining each stage on its 10x cell
barcode, applying transcriptome QC, and retaining a Type-1 geneBC cell or the
reconstructed WT fallback.

```{r reconstruct-final-cells}
# 4. Reconstruct and compare every sample-specific final barcode set
final_stage = c(
  DE_NZ = "DE_GemX",
  DE_XM = "DE",
  ES = "ES",
  PP1 = "PP1",
  PP2 = "PP2",
  S5_1 = "S5_1",
  S5_2 = "S5_2",
  S6_1 = "S6_1",
  S6_2 = "S6_2"
)

stage_validation = lapply(names(final_stage), \(source_stage) {
  tx_path = tx_paths[tx_stages == source_stage]
  bc_path = bc_paths[bc_stages == source_stage]

  stopifnot(length(tx_path) == 1L, length(bc_path) == 1L)

  tx = readRDS(tx_path)
  bc = readRDS(bc_path)

  stopifnot(
    !anyDuplicated(rownames(tx)),
    !anyDuplicated(bc$cellid)
  )

  common_cellids = intersect(rownames(tx), as.character(bc$cellid))
  tx_common = tx[
    match(common_cellids, rownames(tx)),
    ,
    drop = FALSE
  ]
  bc_common = bc[
    match(common_cellids, as.character(bc$cellid)),
    ,
    drop = FALSE
  ]

  stopifnot(
    identical(common_cellids, rownames(tx_common)),
    identical(common_cellids, as.character(bc_common$cellid))
  )

  has_larry_category = (
    !is.na(bc_common$larryBCCategory) &
      bc_common$larryBCCategory != "NA"
  )
  has_gene_category = (
    !is.na(bc_common$geneBCCategory) &
      bc_common$geneBCCategory != "NA"
  )
  is_ko = bc_common$geneBCCategory == type1
  is_wt = (
    bc_common$geneBCCategory == type3 &
      bc_common$larryBCCategory == type1 &
      bc_common$larryBC_type == "WT"
  )
  keep = (
    tx_common$Qualified == "Yes" &
      has_larry_category &
      has_gene_category &
      (is_ko | is_wt)
  )

  reconstructed_ids = paste0(
    sub("-1$", "", common_cellids[keep]),
    "-",
    final_stage[[source_stage]]
  )
  reconstructed_group = ifelse(
    is_ko[keep],
    "KO_cells",
    "WT_cells"
  )

  observed_ids = rownames(final_meta)[
    final_meta$orig.ident == final_stage[[source_stage]]
  ]
  final_index = match(reconstructed_ids, rownames(final_meta))

  summary = data.frame(
    source_stage = source_stage,
    final_label = final_stage[[source_stage]],
    tx_cells = nrow(tx),
    tx_qualified = sum(tx$Qualified == "Yes", na.rm = TRUE),
    bc_cells = nrow(bc),
    joined_cells = length(common_cellids),
    reconstructed = length(reconstructed_ids),
    observed = length(observed_ids),
    reconstructed_only = length(
      setdiff(reconstructed_ids, observed_ids)
    ),
    observed_only = length(setdiff(observed_ids, reconstructed_ids)),
    group_mismatch = sum(
      is.na(final_index) |
        final_meta$group[final_index] != reconstructed_group
    ),
    exact_match = setequal(reconstructed_ids, observed_ids)
  )

  list(summary = summary, ids = reconstructed_ids)
})

validation_summary = do.call(
  rbind,
  lapply(stage_validation, \(x) x$summary)
)
reconstructed_ids = unname(
  unlist(lapply(stage_validation, \(x) x$ids))
)
combined_exact_match = (
  !anyDuplicated(reconstructed_ids) &&
    setequal(reconstructed_ids, rownames(final_meta))
)

knitr::kable(validation_summary, row.names = FALSE)
knitr::kable(
  data.frame(
    check = "Combined reconstructed and final barcode sets are identical",
    reconstructed_cells = length(reconstructed_ids),
    observed_cells = nrow(final_meta),
    passed = combined_exact_match
  ),
  row.names = FALSE
)

stopifnot(
  all(validation_summary$reconstructed_only == 0L),
  all(validation_summary$observed_only == 0L),
  all(validation_summary$group_mismatch == 0L),
  all(validation_summary$exact_match),
  combined_exact_match
)
```

All available checks pass: the documented transcriptome-QC labels are
internally consistent, the recoverable Type-1 UMI thresholds hold, and the
reconstructed final set matches all `r format(nrow(final_meta), big.mark = ",")`
canonical MSK cells exactly.
